## Data Used
The data that I am using for my project is from U.S. Department Of Agriculture.

I obtained the data sets from ers.usda.gov;

The link for the data set named `fruits` is: https://www.ers.usda.gov/data-products/fruit-and-vegetable-prices.aspx 

While the link for the data set named `pTOp` is:https://www.ers.usda.gov/data-products/purchase-to-plate/


The `fruits` and `pTOp` data comes from the Economic Reasearch Service (ERS). Here is a link for more information about the variable names : https://www.ers.usda.gov/data-products/fruit-and-vegetable-prices/documentation/ for `fruits` and https://www.ers.usda.gov/data-products/purchase-to-plate/documentation/ for `pTOp`

The variables in the data include price estimates for over 150 different fruits and estimated prices for food and beverages consumed by participants in "What We Eat America"

I was interested in conducting my own analysis using the data provided to see what is the cheapest stuff to eat, or at the least establish a sort of ranking list

I will be working under the hypothesis that it is cheaper to eat Junk food instead of a balanced diet

## Data Import

This is where I will install any libraries i may need
```{r}
#install.packages("plotly")
library(tidyverse)
library(readxl)
library(esquisse)
library(plotly)
```

Here I'm reading the data sets and assigning them to variables in order to manipulate the data
```{r}
fruits <- read_csv(file="Fruit-prices-2022.csv")
pTOp <- read_csv(file ="pp_national_average_prices_csv.csv")
```

now ill check on the dimensions of the data
```{r}
head(fruits)
dim(fruits)

head(pTOp)
dim(pTOp)
```

Originally the data that comes from the USDA called `fruits` has 62 columns 8 rows.
While the `pTOp` data set contained 15,012 columns and 8 rows aswell


## Data Cleaning/Wrangling

I'm going to label 29 different Types of food from each data set and compare the prices per 100gm of food
```{r}
pTOp <- pTOp %>% mutate(food_type = case_when(
  str_detect(food_description,"Crackers") ~ "Crackers",
  str_detect(food_description,"Yogurt") ~ "Yogurt",
  str_detect(food_description,"buttermilk") ~ "Milk",
  str_detect(food_description,"Milk") ~ "Milk",
  str_detect(food_description, "formula") ~ "Baby food",
  str_detect(food_description, "cocoa") ~ "Chocolate",
  str_detect(food_description, "beef") ~ "Beef",
  str_detect(food_description, "Beef") ~ "Beef",
  str_detect(food_description, "pork") ~ "Pork",
  str_detect(food_description, "ham") ~ "Ham",
  str_detect(food_description, "Lamb") ~ "Lamb",
  str_detect(food_description, "Chicken") ~ "Chicken",
  str_detect(food_description, "Turkey") ~ "Turkey",
  str_detect(food_description, "candy") ~ "Candy",
  str_detect(food_description, "cheese") ~ "Cheese",
  str_detect(food_description, "baby food") ~ "Baby food",
  str_detect(food_description, "candy") ~ "Candy",
  str_detect(food_description, "dressing") ~ "Dressing",
  str_detect(food_description, "Cereal") ~ "Cereal",
  str_detect(food_description, "cereal") ~ "Cereal",
  str_detect(food_description, "Jelly") ~ "Jelly",
  str_detect(food_description, "Jam") ~ "Jam",
  str_detect(food_description, "jam") ~ "jam",
  str_detect(food_description, "jelly") ~ "Jelly",
  str_detect(food_description, "Gelatin") ~ "Gelatin",
  str_detect(food_description, "Candy") ~ "Candy",
  str_detect(food_description, "Fudge") ~ "Fudge",
  str_detect(food_description, "SNICKERS") ~ "Snickers",
  str_detect(food_description, "Baby Ruth") ~ "Baby Ruth",
  str_detect(food_description, "Licorice") ~ "Licorice",
  str_detect(food_description, "MILKY WAY") ~ "Milky Way",
  str_detect(food_description, "Dip") ~ "Dip",
  str_detect(food_description, "dip") ~ "Dip",
  str_detect(food_description, "Ice cream") ~ "Ice Cream",
  str_detect(food_description, "Pudding") ~ "Pudding",
  str_detect(food_description, "chips") ~ "Chips"
  ))

fruits <- fruits %>% mutate(food_type= case_when(
  str_detect(Fruit, "Apple") ~ "Apple",
  str_detect(Fruit, "Apricots") ~ "Apricots",
  str_detect(Fruit, "Bananas") ~ "Bananas",
  str_detect(Fruit, "Berries") ~ "Berries",
  str_detect(Fruit, "Blackberries") ~ "Blackberries",
  str_detect(Fruit, "Cantaloupe") ~ "Cantaloupe",
  str_detect(Fruit, "Cherries") ~ "Cherries",
  str_detect(Fruit, "Clementines") ~ "Clementines",
  str_detect(Fruit, "Cranberries") ~ "Cranberries",
  str_detect(Fruit, "Dates") ~ "Dates",
  str_detect(Fruit, "Fruit cocktail") ~ "Fruit Cocktail",
  str_detect(Fruit, "Grapefruit") ~ "Grapefruit",
  str_detect(Fruit, "Grapes") ~ "Grapes",
  str_detect(Fruit, "Honeydew") ~ "Honeydew",
  str_detect(Fruit, "Kiwi") ~ "Kiwi",
  str_detect(Fruit, "Mangoes") ~ "Mangoes",
  str_detect(Fruit, "Nectarines") ~ "Nectarines",
  str_detect(Fruit, "Oranges") ~ "Oranges",
  str_detect(Fruit, "Papaya") ~ "Papaya",
  str_detect(Fruit, "Peaches") ~ "Peaches",
  str_detect(Fruit, "Pears") ~ "Pears",
  str_detect(Fruit, "Pineapple") ~ "Pineapple",
  str_detect(Fruit, "Plum") ~ "Plum",
  str_detect(Fruit, "Pomegranate") ~ "Pomegranate",
  str_detect(Fruit, "Raspberries") ~ "Raspberries",
  str_detect(Fruit, "Strawberries") ~ "Strawberries",
  str_detect(Fruit, "Watermelon") ~ "Watermelon",
  str_detect(Fruit, "Blueberries") ~ "Blueberries",
  str_detect(Fruit, "Figs") ~ "Figs"
))
```

```{r}
pull(fruits, "food_type") %>% unique() %>% length()

pull(pTOp, "food_type") %>% unique() %>% length()
```
Now that we can see we have two data sets with 29 different food types now its time to clean it up so we only have relevent information
```{r}
fruits29 <- fruits %>% drop_na(food_type)
pTOp29 <- pTOp %>% drop_na(food_type)
head(fruits29)
dim(fruits29)

head(pTOp29)
dim(pTOp29)
```

`fruits29` contains a column named `RetailPrice` which gives the retail price for that food per pound, while `pTOp29` contains a row named `price_100gm` which gives the price for that food per 100 grams.

To get the values on the same scale we will create a new column in `fruits29` that shows the price per 100 grams

```{r}
fruits29 <- fruits29 %>% mutate(price_100gm = RetailPrice*.22)
```

Now we'll clean up the data a little bit so we are only left with relelvent or information that exists in each data set, and then join them.

```{r}
fruits29 <- fruits29 %>% select(Fruit,food_type,price_100gm)
pTOp29 <- pTOp29 %>% select(food_description,food_type,price_100gm)
names(fruits29)[names(fruits29) == 'Fruit'] <- 'food_description'
```

With only the 3 important rows, now we will combine them, and get rid of any messy NAs that resulted in price_100gm, these were due to being not reported in the original `pTOp` dataset

```{r}
doubleSet <- full_join(fruits29,pTOp29)
doubleSet <- doubleSet %>% drop_na(price_100gm)
```

If we check how many unique food types we have then we see we have 4277 observations but only 57 different food types,
```{r}
doubleSet <- arrange(doubleSet,desc(price_100gm),food_type)
pull(doubleSet, "food_type") %>% unique() %>% length()
```

To best represent the data set visually we will take the averages of the different types of foods individually, that way we are left with an average of each type of food that we can compare against each other

```{r}
doubleSet <- doubleSet %>% group_by(food_type) %>% summarize(price = mean(price_100gm))
head(doubleSet)
```
Last clean up step we will give every food a food group, either `Meat`, `Fruit`, `Junk Food`, `Dairy`, or `Baby Food`.
```{r}

doubleSet <- doubleSet %>% mutate(food_group = case_when(
  str_detect(food_type,"Crackers") ~ "Junk Food",
  str_detect(food_type, "Milky Way") ~ "Junk Food",
  str_detect(food_type,"Yogurt") ~ "Dairy",
  str_detect(food_type,"buttermilk") ~ "Dairy",
  str_detect(food_type,"Milk") ~ "Dairy",
  str_detect(food_type, "formula") ~ "Baby Food",
  str_detect(food_type, "Baby food") ~ "Baby Food",
  str_detect(food_type, "Chocolate") ~ "Junk Food",
  str_detect(food_type, "beef") ~ "Meat",
  str_detect(food_type, "Beef") ~ "Meat",
  str_detect(food_type, "Pork") ~ "Meat",
  str_detect(food_type, "Ham") ~ "Meat",
  str_detect(food_type, "Lamb") ~ "Meat",
  str_detect(food_type, "Chicken") ~ "Meat",
  str_detect(food_type, "Turkey") ~ "Meat",
  str_detect(food_type, "candy") ~ "Junk Food",
  str_detect(food_type, "Cheese") ~ "Dairy",
  str_detect(food_type, "baby food") ~ "Baby Food",
  str_detect(food_type, "candy") ~ "Junk Food",
  str_detect(food_type, "Dressing") ~ "Dairy",
  str_detect(food_type, "Cereal") ~ "Junk Food",
  str_detect(food_type, "cereal") ~ "Junk Food",
  str_detect(food_type, "Jelly") ~ "Junk Food",
  str_detect(food_type, "Jam") ~ "Junk Food",
  str_detect(food_type, "jam") ~ "Junk Food",
  str_detect(food_type, "jelly") ~ "Junk Food",
  str_detect(food_type, "Gelatin") ~ "Junk Food",
  str_detect(food_type, "Candy") ~ "Junk Food",
  str_detect(food_type, "Fudge") ~ "Junk Food",
  str_detect(food_type, "Snickers") ~ "Junk Food",
  str_detect(food_type, "Baby Ruth") ~ "Junk Food",
  str_detect(food_type, "Licorice") ~ "Junk Food",
  str_detect(food_type, "Dip") ~ "Junk Food",
  str_detect(food_type, "dip") ~ "Junk Food",
  str_detect(food_type, "Ice Cream") ~ "Junk Food",
  str_detect(food_type, "Pudding") ~ "Junk Food",
  str_detect(food_type, "Chips") ~ "Junk Food",
  str_detect(food_type, "Apple") ~ "Fruit",
  str_detect(food_type, "Apricots") ~ "Fruit",
  str_detect(food_type, "Bananas") ~ "Fruit",
  str_detect(food_type, "Berries") ~ "Fruit",
  str_detect(food_type, "Blackberries") ~ "Fruit",
  str_detect(food_type, "Cantaloupe") ~ "Fruit",
  str_detect(food_type, "Cherries") ~ "Fruit",
  str_detect(food_type, "Clementines") ~ "Fruit",
  str_detect(food_type, "Cranberries") ~ "Fruit",
  str_detect(food_type, "Dates") ~ "Fruit",
  str_detect(food_type, "Fruit Cocktail") ~ "Fruit",
  str_detect(food_type, "Grapefruit") ~ "Fruit",
  str_detect(food_type, "Grapes") ~ "Fruit",
  str_detect(food_type, "Honeydew") ~ "Fruit",
  str_detect(food_type, "Kiwi") ~ "Fruit",
  str_detect(food_type, "Mangoes") ~ "Fruit",
  str_detect(food_type, "Nectarines") ~ "Fruit",
  str_detect(food_type, "Oranges") ~ "Fruit",
  str_detect(food_type, "Papaya") ~ "Fruit",
  str_detect(food_type, "Peaches") ~ "Fruit",
  str_detect(food_type, "Pears") ~ "Fruit",
  str_detect(food_type, "Pineapple") ~ "Fruit",
  str_detect(food_type, "Plum") ~ "Fruit",
  str_detect(food_type, "Pomegranate") ~ "Fruit",
  str_detect(food_type, "Raspberries") ~ "Fruit",
  str_detect(food_type, "Strawberries") ~ "Fruit",
  str_detect(food_type, "Watermelon") ~ "Fruit",
  str_detect(food_type, "Blueberries") ~ "Fruit",
  str_detect(food_type, "Figs") ~ "Fruit"))

doubleSetavg <- doubleSet %>% group_by(food_group) %>% summarize(price = mean(price))
```
## Data Visualization

This first graph shows all the food groups plotted next to each other, while the second graph shows the food groups plotted with like food groups, making it easier to see the differences in prices among different types of foods in the same food group.


This graph, named `doubleGraph1`, shows us that `Lamb` is the most expensive food among all the food groups
```{r}

doubleGraph1 <- ggplot(doubleSet) +
  aes(
    x = price,
    y = food_type,
    fill = food_group,
    group = food_group
  ) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Price",
    y = "Food Name",
    title = "Food Price",
    subtitle = "(On Average,
    Per 100 Gram)"
  ) +
  theme_minimal()+
  scale_fill_discrete(name = "Food Groups")

ggplotly(doubleGraph1)
```


The second graph shows us that `Snickers` is the most expensive `junk food`, `Raspberries`, the most expensive `Fruit`, `Cheese` the most expensive `Dairy`, and `Lamb` is the most expensive `Meat`

```{r}
doubleGraph2 <-ggplot(doubleSet) +
  aes(x = price, y = food_type, fill = food_group) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_hue(direction = 1) +
  labs(x = "Price", y = "Food Name", title = "Food Prices") +
  theme_minimal() +
  facet_wrap(vars(food_group))+
  scale_fill_discrete(name = "Food Groups")

ggplotly(doubleGraph2)
```


Then I got to thinking, what about whats the most expensive of the 5 groups, so I made one more graph

```{r}
averageGraph <- ggplot(doubleSetavg) +
  aes(x = price, y = food_group, fill = food_group) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Price",
    y = "Name",
    title = "Most Expensive Food Groups"
  ) +
  theme_minimal()

ggplotly(averageGraph)
```

I came to the conclusion that i need more data to better represent all the food groups, but it would seem that my working hypothesis was false, it would be cheaper to eat fruits and dairy it seems.