## Data Used
The data that I am using for my project is from U.S. department of Agriculture.

I obtained the data sets from ers.usda.gov;

The link for the data set named `fruits` is: https://www.ers.usda.gov/data-products/fruit-and-vegetable-prices.aspx 

While the link for the data set named `pTOp` is:https://www.ers.usda.gov/data-products/purchase-to-plate/


The `fruits` and `pTOp` data comes from the Economic Reasearch Service (ERS). Here is a link for more information about the variable names : https://www.ers.usda.gov/data-products/fruit-and-vegetable-prices/documentation/ for `fruits` and https://www.ers.usda.gov/data-products/purchase-to-plate/documentation/ for `pTOp`

The variables in the data include price estimates for over 150 different fruits and estimated prices for food and beverages consumed by participants in "What We Eat America"

I was intersted in conducting my own analysis using the data provided to see what is the cheapest stuff to eat, or at the least establish a sort of ranking list





This is where I will install any librarys i may need
```{r}
library(tidyverse)
library(readxl)
```

## Data Import

Here i'll b reading the data
```{r}
fruits <- read_csv(file="Fruit-prices-2022.csv")

pTOp <- read_csv(file ="pp_national_average_prices_csv.csv")

```

now ill check on the dimensions of the data
```{r}
head(fruits)
dim(fruits)

head(pTOp)
dim(pTOp)


```
here ill be adding thigns to the data that make it easier to grab latter on.
```{r}
pTOp <- pTOp %>% mutate(food_type = case_when(
  str_detect(food_description,"milk") ~ "Milk",
  str_detect(food_description,"yogurt") ~ "yogurt",
  str_detect(food_description,"buttermilk") ~ "Milk",
  str_detect(food_description,"Milk") ~ "Milk",
  str_detect(food_description, "formula") ~ "Baby food",
  str_detect(food_description, "cocoa") ~ "Chocolate",
  str_detect(food_description, "cream") ~ "Creamer",
  str_detect(food_description, "beef") ~ "beef",
  str_detect(food_description, "Beef") ~ "beef",
  str_detect(food_description, "pork") ~ "Pork",
  str_detect(food_description, "ham") ~ "Ham",
  str_detect(food_description, "Lamb") ~ "Lamb",
  str_detect(food_description, "Chicken") ~ "Chicken",
  str_detect(food_description, "Turkey") ~ "Turkey"
  ))
```

```{r}

```

```{r}

```